The paper on \Large \bf Blockchain Frontrunning

### 2018-09-07 frontrunning/Latex/sections/paper.tex
[frontrunning] some notes

% !TEX root = ../main.tex
%______________________________________________________________%
\section{Introduction}

@@

%______________________________________________________________%
\section{Preliminaries}

@@

\emph{Front-running} is analogous to any course of action during which a person benefits from prior access to inside or confidential market information about upcoming transactions and trades.  The idea of front-running was originated back in stock markets. In the old times, all the trades were executed on papers, so in case of receiving a large order from a client, the broker might say this loud and other people around the table could be informed. Therefore, a malicious trader would now run in front of that order and put his own transaction in-between (before the trade was executed) and profit from the price increase of the stock. This problem occurs when a group of market participants obtain non-public market information which allows them to front-run other users trades by putting their orders a head of those trades and benefit from advance knowledge of pending client orders
Nowadays, there are a wide range of decentral applications implemented using blockchain technology. Being decentralized, there is no central version of information to be attacked, however, due to the inherent nature of the blockchain infrastructure, the data flow is transparent hence these applications are open to \emph{front-running} attacks. One such application is \textbf{Bancor} -- an Ethereum-based application that allows users to exchange their tokens without any counter-party risk. Researchers have shown that this financial blockchain application is open to a simple front-running attack.  However, front-running vulnerability is not limited to the financial application and it can happen in other types of blockchain applications as well.

@@

%___________________________________________________________%
\section{Related Work}
As the traditional front-running was originated to trading financial instruments most of the literature are focused on financial aspects of the markets. %TODO: expand this with something

@@

@@

% More to read:  SECURITIES AND EXCHANGE COMMISSION Release No. 34-67079. pdf included in Related_Documents
% https://www.sec.gov/about/annual_report/1988.pdf

On the years after there were ongoing discussions between self-regulatory exchanges (\eg \textit{CBoE}) and  \textit{SEC} to regulate, detect and define laws and regulations to deal with front-running~\cite{markham1988front}, with \textit{SEC} stating: 
\begin{quote}
It seems evident that such behavior on the part of persons with knowledge of imminent transactions which will likely affect the price of the derivative security constitutes an unfair use of such knowledge. \footnote{Securities Exchange Act Release No. 14156, November 19, 1977, (Letter from George A. Fitzsimmons, Secretary, Securities and Exchange Commission to Joseph W. Sullivan, President  CBoE).}
\end{quote}

As defining what exactly is considered illegal front-running required more knowledge of how these new markets behave, \textit{CBoE} and other exchanges (and brokers) issued educational circulars for their members asserting that front-running violates existing rules, with some examples of what is considered front-running. However difference of opinion regarding the unfairness of front running activities, insufficient exchange rules and lack of a precise definition in this area resulted in no action by self-regulator organizations~cite{sec1978optionsmarket}.

Further reading on the early details on the history and challenges of detecting and regulating front-running can be found in~\cite{markham1988front} . %add a more recent publication here as well!

@@

Ethereum blockchain contains all transaction ever made on Ethereum, however, it is not easy to see the behaviour of the participants and miners solely using the default clients nor online blockchain explorers. In order to get the required data, Ethereum blockchain data should be converted to a queryable format. For this case study, we used open source projects such as Go Ethereum implementation\footnote{Official Go implementation of the Ethereum protocol\url{https://github.com/ethereum/go-ethereum}}, python script for extracting, transforming and loading ethereum blocks, named \texttt{ethereum-etl}~\cite{ethereumetl} and Google BigQuery\footnote{A fast, highly scalable, cost-effective, and fully managed cloud data warehouse for analytics \url{https://cloud.google.com/bigquery/}}. Using this software stack, we were able to query for all the transactions and blocks within the window of Status.im ICO for further analysis which was mainly done by data analysis tool \texttt{Tableau}\footnote{Tableau is a data visualization software that is used for data science and business intelligence.\url{https://www.tableau.com/}}. A copy of this dataset and the findings can be found in \texttt{github.com / ANONYMOUS}.

@@

Although this incident does not involve transaction reordering in the blocks, it shows how miners can modify their code to behave in a certain way to front-run other transactions and take advantage of the monetary profit. The dataset used for this analysis can be found on our GitHub repository.\footnote{Frontrunning Status.im ICO dataset: \url{http://github.com/TODO:ANONYMOUS}}. This analysis made us wonder, how many other DApps are vulnerable to front-running attack, who can front-run and what are the mitigation methods for decentral applications?

As the result of this analysis, we discovered a wide range of blockchain front-running case studies that are yet to investigated. Therefore, in the next sections we dive deeper into blockchain specific front-running details and analyze top recent DApps for front-running attack vectors and possible solutions proposed to mitigate these issues.

@@

\subsubsection{Miners}
Blockchain miners are the only parties who can decide on the order of transactions within a block they mine, they can easily intercept and reorder the transactions in their blocks, this is known as \textbf{transaction reordering} attack.  For example in an Ethereum based application, a miner learns about the pending \textit{buy} transaction of 1000 units of a token\ie TKN, presumably if this transaction goes through, it causes the price of the purchased token to increase. So the dishonest miner can step in front of this transaction and place his own buy order ahead. He would simply create his \textit{buy} of 1000 TKN and include it in front of the previous \textit{buy} transaction of 1000 TKN in the block he mines. Doing so, he would receive a better rate than other network users, and afterwards can sell the assets he has bought and gain a price advantage at the expense of others. Similarly, a dishonest miner can sell his tokens in front, if he sees a pending \textit{sell} transaction. In an open source blockchain application, there is no rules on how transactions must be ordered and miners are free to mine transactions in any order they prefer. However, miners can only front-run transactions within the blocks they happen to mine. They could do this also by not broadcasting their own transactions to the network, this makes the miners to be less noticeable to the network when front-running.
Another example here is in the case of decentral exchanges and order cancellation. Every exchange should have the functionality to cancel the orders, especially for a volatile market. In this case, when user decides that her order is not profitable anymore, she would send a cancelation transaction. Here the miner sees the cancelation transaction and puts his own order in front of the cancellation transaction --transaction reordering-- to fill the order, potentially profiting from the order. They can also include the canceled transaction in their block to collect the full gas limit used by that transaction~\cite{CostofDecentralization:online}, see Figure~\ref{MinerFrontrunning}. Front-running can also occur in non-financial smart contracts. As an example, smart contract which adds all the participants to a party invite list, could only be closed by the smart contract owner. In this case when the owner sends the transaction to close the list, a miner can include his own list of participants before the close transaction in the block he mines. This is an issue with the design of the smart contracts and is known as transaction-ordering dependence vulnerability~\cite{luu2016making}.

@@

\subsection{Case Studies}

@@

%Some edits required here! @mahsa TODO

@@

%______________________________________________________________%
\subsection{Crypto-Collectibles Games} \hfill\\

%TODO: shayan: connect games to cryptokitties and other collectibles

Crypto-Collectibles, also known as ERC-721 Non-Fungible Token Standard~\cite{erc721}, are one of the applications of Ethereum blockchain. Cryptokitties is known to be one of the first and the most popular use cases of ERC721 tokens~\cite{cryptokitties}. Their marketcap reached more than 6 million dollars in the first months their launch. Crypto-Collectibles, are similar to ERC-20 tokens and can be listed on exchanges. Most of the front-running issues applied to the exchanges also apply to DApps using ERC-721 tokens. \\

As an example Cryptokitties can be bred, bought or sold through the main interface\footnote{Cryptokitties website \url{https://www.cryptokitties.co/}}. See Figure~\ref{cryptokittie}.
To buy a cryptokittie, user sends a bid transaction \textit{bid(uint256 \_tokenId)} containing the cryptokittties ID which can be found either on their website or on the ERC721 token smart contract. This is similar to open auctions, bid value and the object bidding on is visible to the network and any user could be easily front-run by sending the same transaction with higher gasPrice to replace the initial bidder. 
% https://etherscan.io/tx/0x62a76b7e345e8abd463487d0348c118a9314b197024e090fee7b1fbce434da40

@@

%TODO: Shayan: Frontrunning on one of the gambling DAPPS from the table

%______________________________________________________________%
\section{How to stop Front-running?} %We have have the applicability of the solutions inline.

@@

\emph{Traditional Front-running Prevention Methods}. regulations/enforcement/broker education/sealed order
%@shayan: ^ TODO!

@@

There are two main approaches to prevent front running, one to design a blockchain that is front-run resistant , and the other to design the application logics in a way that front running is not possible. 
%@shayan: ^ TODO!

%______________________________________________________________%
\subsection{Front-run Resistant Blockchain}
What does this mean to have a front-run resistant blockchain?  There are technical difficulties to achieve such solutions as there are unknown factors within such network designs (corner/edge cases). In this section, we describe the potential solutions using which, one can design and implement a decentral application that is resistant to front-running.

@@

Market participants are concerned about revealing information on their past trades, purchases, income, or liquidity needs. Privacy is a basic human right, that being said, there are limitations with the functionality of a system designed to preserve full privacy. However other than past transactions, real time transactions could leak informations about the sender or the intention of the order that could lead to front-running attacks. Privacy preserving blockchains try to solve the initial issue by keeping all the details of the transactions private, in some extent~\cite{miller2017empirical, kappos2018empirical}. 
As an example, ZCash~\cite{hopwood2016zcash}, uses two distinct type of addresses, transparent addresses and shielded addresses. Transparent addresses work similar to Bitcoin transactions, fully transparent about the sender and receiver addresses, the amount and included data. However shielded addresses are private and do not leak any information, this is great for privacy but not good for smart contracts capabilities. Smart contract functionality requires verification by the miners which require to know the input, functionality and output of smart contracts to be able to determine if the transaction is valid. Albeit by using cryptographic primitives such as zero-knowledge proofs this issue could also be resolved~\cite{kosba2016hawk}.

\subsubsection{Design Decision \#3: Loopring/Dual Authoring\newline}
Talk about this possible solution
%MAHSA TODO
%https://medium.com/loopring-protocol/dual-authoring-looprings-solution-to-front-running-d0fc9c348ef1
% https://raw.githubusercontent.com/Loopring/whitepaper/master/en_whitepaper.pdf

@@

%______________________________________________________________%
\subsection{Application Design to Prevent Front-running}

@@

%______________________________________________________________%
\section{Concluding Remarks}

>>>>

% !TEX root = ../main.tex

\section{Introduction}

@@

\section{Preliminaries}

@@

\emph{Front-running} is analogous to any course of action during which a person benefits from prior access to inside or confidential market information about upcoming transactions and trades. This problem can occur in both financial and non-financial systems, however, it is more noticeable within the trading and exchange systems as it was originated back in stock markets. In the old times, all the trades were executed on papers, so in case of receiving a large order from a client, the broker might say this loud and other people around the table could be informed. Therefore, a malicious trader would now run in front of that order and put his own transaction in-between (before the trade was executed) and profit from the price increase of the stock. In other words, a group of market participants obtain non-public market information which allows them to front-run other users trades by putting their orders a head of those trades and benefit from advance knowledge of pending client orders. The two significant factors which cause the front-running practice to happen within the financial markets are (i) imperfect competition and (ii) liquidity uncertainty~\cite{liang2005distressed}. Any sort of front-running activity within the financial markets is considered as an unethical and illegal practice as it is unfairly beneficial for a few market participants who have the privilege of acting on this information and taking advantages at the expense of the investors.

@@

%___________________________________________________________%

\section{Related Work}
As the traditional front-running was originated to trading financial instruments most of the literature are focused on financial aspects of the markets. %TODO: expand this with something

@@

% talk about the AMF frontrunning.

@@

%In actuality, front-running is more complex than this definition suggests. It encompasses at least three forms of conduct, each of which raises different regulatory and policy issues.9 They are: (1) trading by third parties who are tipped on an impending block trade ("tippee" trading); (2) transactions in which the owner or purchaser of the block trade itself engages in the offset- ting futures or options transaction as a means of "hedging" against price fluctuations caused by the block transaction ("self-front-running"); 1 ° and (3) transactions where a broker with knowledge of an impending customer block order trades ahead of that order for the broker's own profit ("trading ahead").
% from JerryWMarkhamfront-running.pdf

% More to read and add from : SECURITIES AND EXCHANGE COMMISSION Release No. 34-67079. pdf included in Related_Documents
% https://www.sec.gov/about/annual_report/1988.pdf

On the years after there were ongoing discussions between self-regulatory exchanges (\eg \textit{CBoE}) and  \textit{SEC} to regulate, detect and define laws and regulations to deal with front-running~\cite{markham1988front}, with \textit{SEC} stating: 
\begin{quote}
It seems evident that such behaviour on the part of persons with knowledge of imminent transactions which will likely affect the price of the derivative security constitutes an unfair use of such knowledge. \footnote{Securities Exchange Act Release No. 14156, November 19, 1977, (Letter from George A. Fitzsimmons, Secretary, Securities and Exchange Commission to Joseph W. Sullivan, President  CBoE).}
\end{quote}

As defining what exactly is considered illegal front-running required more knowledge of how these new markets behave, \textit{CBoE} and other exchanges (and brokers) issued educational circulars for their members asserting that front-running violates existing rules, with some examples of what is considered front-running. However difference of opinion regarding the unfairness o front running activities, insufficient exchange rules and lack of a precise definition in this area resulted in no action by self-regulator organizations~cite{sec1978optionsmarket}.

Further reading on the early details on the history and challenges of detecting and regulating front-running can be found in~\cite{markham1988front} . % TODO: add a more recent publication here as well!

@@

@@

Although this incident does not involve transaction reordering in the blocks, it shows how miners can modify their code to behave in a certain way to front-run other transactions and take advantage of the monetary profit. The dataset used for this analysis can be found on our GitHub repository.\footnote{Frontrunning Status.im ICO dataset: \url{http://github.com/TODO:ANONYMOUS}}
%TODO: these might need more details and proofs. ask Jeremy about how to go on about this!

@@

\subsubsection{Miners}
Blockchain miners are the only parties who can decide on the order of transactions within a block they mine, they can easily intercept and reorder the transactions in their blocks, this is known as \textbf{transaction reordering} attack.  For example in an Ethereum based application, a miner learns about the pending \textit{buy} transaction of 1000 units of a token, i.e. TKN, presumably if this transaction goes through, it causes the price of the purchased token to increase. So the dishonest miner can step in front of this transaction and place his own buy order ahead. He would simply create his \textit{buy} of 1000 TKN and include it in front of the previous \textit{buy} transaction of 1000 TKN in the block he mines. Doing so, he would receive a better rate than other network users, and afterwards can sell the assets he has bought and gain a price advantage at the expense of others. Similarly, a dishonest miner can sell his tokens in front, if he sees a pending \textit{sell} transaction. In an open source blockchain application, there is no rules on how transactions must be ordered and miners are free to mine transactions in any order they prefer. However, miners can only front-run transactions within the blocks they happen to mine. They could do this also by not broadcasting their own transactions to the network, this makes the miners to be less noticeable to the network when front-running.
Another example here is in the case of decentral exchanges and order cancellation. Every exchange should have the functionality to cancel the orders, especially for a volatile market. In this case, when user decides that her order is not profitable anymore, she would send a cancelation transaction. Here the miner sees the cancelation transaction and puts his own order in front of the cancellation transaction --transaction reordering-- to fill the order, potentially profiting from the order. They can also include the canceled transaction in their block to collect the full gas limit used by that transaction~\cite{CostofDecentralization:online}, see Figure~\ref{MinerFrontrunning}. Front-running can also occur in non-financial smart contracts. As an example, smart contract which adds all the participants to a party invite list, could only be closed by the smart contract owner. In this case when the owner sends the transaction to close the list, a miner can include his own list of participants before the close transaction in the block he mines. This is an issue with the design of the smart contracts and is known as transaction-ordering dependence vulnerability~\cite{luu2016making}.

@@

\subsection{OCase Studies}

@@

%Some edits required here!

@@

\subsection{Crypto-Collectibles auctions} \hfill\\ 
Crypto-Collectibles, also known as ERC-721 Non-Fungible Token Standard~\cite{erc721}, are one of the applications of Ethereum blockchain. Cryptokitties is known to be the most popular applications of ERC721 tokens~\cite{cryptokitties}. The main usage of these Cryptokitties is that users can breed them with each other and make new unique Cryptokitties, also they can buy or sell them through their main interface\footnote{Cryptokitties website \url{https://www.cryptokitties.co/}}. See Figure~\ref{cryptokittie}.
To buy a cryptokittie, user sends a bid transaction \textit{bid(uint256 \_tokenId)} containing the cryptokittties ID which can be found either on their website or on the ERC721 token smart contract. This is similar to open auctions and everyone can see the bid value and what is the bid for, this transaction is visible from any other Ethereum node and could be easily front-run by sending the same transaction with higher gasPrice. 
%TODO: this needs a rewrite. 
% https://etherscan.io/tx/0x62a76b7e345e8abd463487d0348c118a9314b197024e090fee7b1fbce434da40

@@

\section{How to stop Front-running?} %We have have the applicability of the solutions inline.

@@

\emph{Traditional Front-running Prevention Methods}. regulations/enforcement/broker education/sealed order
%@shayan: ^

@@

There are two main approaches to prevent front running, one to design a blockchain that is front-run resistant , and the other to design the application logics in a way that front running is not possible.

\subsection{Front-run Resistant Blockchain}
What does this mean to have a front-run resistant blockchain?  There are technical difficulties to achieve such solutions as there are unknown factors within such network designs (corner/edge cases). In this section, we describe the potential solutions using which, one can design and implement a decentral application that is resistant to front-running.

@@

%NEEDS TO BE DONE:
Market participants are concerned about revealing information on their past trades, purchases, income, or liquidity needs (forget about market participants >> this leads to frontrunning)
Shielded transactions in ZCash do not reveal the sender, receiver, the amount and the data included in the transactions, hence they cannot be seen by network participants to be front-run. Although this limits the functionality of the blockchain/smartcontracts/dapps

\subsubsection{Design Decision \#3: Loopring/Dual Authoring\newline}
Talk about this possible solution

@@

\subsection{Application Design to Prevent Front-running}

@@

\section{Concluding Remarks}
